### 2018-11-12 source/Ch8-hiresEGS/hires-EGS.tex
asdasda

\begin{table}[h!]
	\begin{tabular}{ccccc}
		\# & RA [hms]    & Dec [dms]   & Dist. from EGS [deg.] & Dist. from 3C295 [deg.] \\\hline
		1  & 14:30:18.72 & 52:17:29.80 & 2.041                         & 2.904 \\
		2  & 14:19:44.44 & 54:23:04.58 & 1.928                         & 2.517 \\ 
		3  & 14:21:20.05 & 53:03:46.00 & 0.864                         & 1.743 \\
		4  & 14:21:09.41 & 51:22:32.46 & 1.294                         & 1.728 \\
		5  & 14:11:50.32 & 52:49:02.66 & 0.844                         & 0.619 \\
		6  & 14:11:20.23 & 52:12:04.30 & 0.915                         & 0.000 \\
		7  & 14:08:07.00 & 52:55:11.36 & 1.409                         & 0.869 \\
		8  & 14:08:09.76 & 52:44:46.56 & 1.354                         & 0.680 \\
	\end{tabular}
	\caption{\label{table.LOBOS.sources}Table giving the positions of our 8 LBCS sources and their distance from both the centre of the EGS and 3C295.}
\end{table}
\begin{figure}[h!]
	\includegraphics[width=\linewidth]{images/{image_full_ampphase_di_m.NS_Smooth.noise01.fitslbcs.egs.positions}.png}
	\caption{Position of the calibrator sources around the EGS, overlaid on the widefield EGS image. LBCS is 3C295.}
	\label{lbcs-coverage-image}
\end{figure}

\pg
A significant preliminary test is therefore to check that we do in fact see any of the LBCS sources. However, this is not the end of the problem: decorrelation causes flux from a source to be ``smeared" (hence its other name) around the source. The DDFacet suite can account for this effect by modelling the direction-dependent PSF, i.e. a PSF which is smeared appropriately as a function of position\footnote{A single smeared PSF is calculated per facet, and this smeared PSF is treated as being constant within the facet: this is a good approximation unless one uses truly massive facets.}.

@@

\pg
We select 8 LBCS sources, which are all the ``good" sources within 5 degrees of the centre of the EGS. \cref{lbcs-coverage-image} shows the position of these calibrator sources with respect to the EGS. If we compare \cref{lbcs-coverage-image} and \cref{plot.EGS.lofar.widefield}, we can immediately see that the EGS lies in two separate facets of our direction-dependent calibration, which are the eastern neighours of the 3C295 facet. Neither of these EGS facets contains a LBCS source (indeed, if they did, our work would have perhaps been easier); however, nearby facets do contain one or multiple LBCS calibrators. If the effect of directional gain errors for these sources is comparable to those for the EGS facets, then we can estimate their impact on an eventual high-resolution image of the EGS.

@@

\pg
The LBCS sources were chosen automatically using a routine written by Leah Morabitoh and shared in a personal communication. This was done in early 2017. Only the brightest reliable sources were kept, provided they had a good reliability factor (the LBCS catalog includes a reliability factor for each source; we have not verified that these reliability factors have not changed since). This left a total of 8 sources, described in \cref{table.LOBOS.sources}. Their position on the sky is shown in \cref{lbcs-coverage-image}. They will henceforth be referred to as LBCS1, LBCS2, and so on - note that LBCS6 is actually 3C295.

@@

\pg
Finally, because 3C295 is such a bright source, the dynamic range around it is extremely high and so the overlay minimum lies at $100\sigma$. We only create a single overlay for this source (as opposed to 2 for all others), because it is already discussed in detail in \cref{section.3c295}. The associated overlay is shown in \cref{fig.lbcs6.100sigma.overlay}: we see that the low-resolution image shows roughly the same morphology (e.g. northern/southern lobe directions), albeit at much lower resolution. The other overlays are shown below.
\begin{figure}[h!]
\includegraphics[width=0.75\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs6.overlay.100sigma}.png}
\caption{Overlay of international LOFAR image of 3C295 over the widefield image.}
\label{fig.lbcs6.100sigma.overlay}
\end{figure}

@@

\pg
A few things are immediately apparent: LBCS1 seems to be a binary source, and LBCS3 seems to suffer strongly from directional gain errors (the source structure in stripes seems unlikely to be physical). LBCS4 seems almost unaffected, which would indicate that the portions of the EGS which lie between 3C295 and LBCS4 are not strongly affected by directional gain errors. The fact that LBCS3 seems to be affected would corroborate the notion that the characteristic scale of DDE impacts is of the order of just over a degree.%, LBCS5, and LBCS7, however, show the presence of very strong artefacts, even at the $15\sigma$ level. In the case of LBCS8, the absence of artefacts is likely due to the fact that the source itself is extremely faint.

\pg
Of course, it is not enough to eyeball this: we therefore proceed to quantify the impact of directional gain errors for different sources in the field. To do this, we begin by making the approximation that they are a function of distance from the calibrator source only, rather than specifically a function of direction and distance. This is simply because we don't have enough calibrators to do otherwise. Secondly, we treat the effect of directional gain errors as \textit{increasing the noise} around a source compared to what the thermal noise is expected to be around it: this thermal noise is expected to increase as a function of distance from the phase centre because we apply a beam-correction during imaging (i.e. because we do not image apparent flux but integrated flux). \cref{fig.noiserms.distEGS} shows this effect in practice; note that the noise estimation around LBCS6 (3C295) is biased because we are still quite close to this extremely bright source, and thus likely have a large contribution from artefacts. This explains its outlier behaviour.
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/RMSvsDistFromEGS.png}
\caption{Thermal noise near LBCS sources as a function of distance from the phase centre.}
\label{fig.noiserms.distEGS}
\end{figure}

\pg
To estimate the prevalence of directional gain errors, we then calculate the rms around each LBCS source. This is plotted in \cref{fig.dderms.distEGS}. The RMS around each source is normalised by the source's integrated flux as calculated on the restored image: this is because we expect brighter images to show stronger artefacts associated to directional gain errors. It should be noted that LBCS8 is not deconvolved at all: as such, its RMS is effectively calculated on a corrupted dirty map of the source, rather than a residual map. This helps explains its outlier behaviour. This outlier aside, the directional gain rms seems fairly flat as a function of distance from 3C295, however - a very good sign, as it indicates that the directional gain errors are comparable across the primary beam. However, we do not take into account the increase in thermal rms as a function of distance from phase centre in this plot. We therefore look at the normalised RMS values for each LBCS source, shown in \cref{fig.rmsratio.distEGS}. The relationship remains flat, though LBCS5 now stands as a $3\sigma$ outlier. The blue line shows the mean value, excluding LBCS8; it lies at $y=1.08063981$. In other words, the average RMS induced by directional gain errors, when normalised by both the flux of a source for which it is calculated and the thermal noise, is very close to unity, and does not increase very much as distance from 3C295 increases. While it apparently increases in the direction of LBCS5, the EGS is not in this direction from 3C295, and so this is not a source of concern for our project of imaging the EGS using international LOFAR.
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/ArtefactRMSvsDistFrom3c295.png}
\caption{DDE noise near LBCS sources as a function of distance from the phase centre. Because 3C295 is so much brighter than all other sources ($\sim100$Jy; second-brightest LBCS source is $\sim2$Jy), its DDE noise was normalised by its integrated model flux.}
\label{fig.dderms.distEGS}
\end{figure}
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/NormArtefactRMSvsDistFrom3c295.png}
\caption{Normalised RMS values around each of the LBCS sources in the field as a function of distance from phase centre. }
\label{fig.rmsratio.distEGS}
\end{figure}

\pg
The most encouraging result shown above, however, is for LBCS3 \& LBCS4: both of these sources appear to be only very slightly affected by directional gain errors. Since the EGS lies between these two calibrator sources and 3C295, this is extremely positive. Similarly, the fact that LBCS4 resolves into two separate sources (as does LBCS1, for that matter) would indicate that even neighbouring bright sources do resolve properly - this is not a rigorous test, but it is a fortunate serendipitous observation to be able to make.

>>>>

\begin{table}[h!]
	\begin{tabular}{ccccc}
		\# & RA [hms]    & Dec [dms]   & Dist. from EGS [deg.] & Dist. from 3C295 [deg.] \\\hline
		1  & 14:30:18.72 & 52:17:29.80 & 2.041                         & 2.904 \\
		2  & 14:19:44.44 & 54:23:04.58 & 1.928                         & 2.517 \\ 
		3  & 14:21:20.05 & 53:03:46.00 & 0.864                         & 1.743 \\
		4  & 14:21:09.41 & 51:22:32.46 & 1.294                         & 1.728 \\
		5  & 14:11:50.32 & 52:49:02.66 & 0.844                         & 0.619 \\
		6  & 14:11:20.23 & 52:12:04.30 & 0.915                         & 0.000 \\
		7  & 14:08:07.00 & 52:55:11.36 & 1.409                         & 0.869 \\
		8  & 14:08:09.76 & 52:44:46.56 & 1.354                         & 0.680 \\
	\end{tabular}
	\caption{\label{table.LOBOS.sources}Table giving the positions of our 8 LBCS sources and their distance from both the centre of the EGS and 3C295.}
\end{table}
\begin{figure}[h!]
	\includegraphics[width=\linewidth]{images/{image_full_ampphase_di_m.NS_Smooth.noise01.fitslbcs.egs.positions}.png}
	\caption{Position of the calibrator sources around the EGS, overlaid on the widefield EGS image. LBCS6 is 3C295.}
	\label{lbcs-coverage-image}
\end{figure}

\pg
A significant preliminary test is therefore to check that we do in fact see any of the LBCS sources. However, this is not the end of the problem: decorrelation causes flux from a source to be ``smeared" (hence its other name) around the source. The DDFacet suite can account for this effect by modelling the direction-dependent PSF, i.e. a PSF which is smeared appropriately as a function of position\footnote{A single smeared PSF is calculated per facet, and this smeared PSF is treated as being constant within the facet: this is a good approximation unless one uses truly massive facets.}. The effect of this smearing is shown in \cref{image.ddPSF}.

@@

\pg
We select 8 LBCS sources, which are all the ``good" sources within 5 degrees of the centre of the EGS. \cref{lbcs-coverage-image} shows the position of these calibrator sources with respect to the EGS. If we compare \cref{lbcs-coverage-image} and \cref{plot.EGS.lofar.widefield}, which are the same image at different levels of contrast, we can immediately see that the EGS lies in two separate facets of the direction-dependent calibration we performed without the international baselines: these are the eastern neighours of the 3C295 facet. Neither of these EGS facets contains a LBCS source (indeed, if they did, our work would have perhaps been easier); however, nearby facets do contain one or multiple LBCS calibrators. If the effect of directional gain errors for these sources is comparable to those for the EGS facets, then we can estimate their impact on an eventual high-resolution image of the EGS.

@@

\pg
The LBCS sources were chosen automatically using a routine written by Leah Morabitoh and shared in a personal communication. This was done in early 2017. Only the brightest reliable sources were kept, provided they had a good reliability factor (the LBCS catalog includes a reliability factor for each source; we have not verified that these reliability factors have not changed since). We thus had 8 LBCS sources around the field, described in \cref{table.LOBOS.sources}. Their position on the sky is shown in \cref{lbcs-coverage-image}. They will henceforth be referred to as LBCS1, LBCS2, and so on - note that LBCS6 is actually 3C295.

@@

\pg
Finally, because 3C295 is such a bright source, the dynamic range around it is extremely high and so the overlay minimum lies at $100\sigma$. We only create a single overlay for this source (as opposed to 2 for all others), because it is already discussed in detail in \cref{section.3c295}. The associated overlay is shown in \cref{fig.lbcs6.100sigma.overlay}: we see that the low-resolution image shows roughly the same morphology (e.g. northern/southern lobe directions), albeit at much lower resolution. The other overlays are shown in \cref{fig.lbcs1.overlay.5sigma,fig.lbcs2.overlay.5sigma,fig.lbcs3.overlay.5sigma,fig.lbcs4.overlay.5sigma,fig.lbcs5.overlay.5sigma,fig.lbcs6.overlay.5sigma,fig.lbcs7.overlay.5sigma,fig.lbcs8.overlay.5sigma}.
\begin{figure}[h!]
\includegraphics[width=0.75\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs6.overlay.100sigma}.png}
\caption{Overlay of international LOFAR image of 3C295 over the widefield image.}
\label{fig.lbcs6.100sigma.overlay}
\end{figure}

@@

\pg
A few things are immediately apparent: LBCS1 (\cref{fig.lbcs1.overlay.5sigma}) seems to be a binary source, and LBCS3 (\cref{fig.lbcs3.overlay.5sigma}) seems to suffer strongly from directional gain errors (the source structure in stripes seems unlikely to be physical). LBCS4 (\cref{fig.lbcs4.overlay.5sigma}) seems almost unaffected, which would indicate that the portions of the EGS which lie between 3C295 and LBCS4 are not strongly affected by directional gain errors. The fact that LBCS3 seems to be affected would corroborate the notion that the characteristic scale of DDE impacts is of the order of just over a degree.%, LBCS5, and LBCS7, however, show the presence of very strong artefacts, even at the $15\sigma$ level. In the case of LBCS8, the absence of artefacts is likely due to the fact that the source itself is extremely faint.

\pg
Of course, it is not enough to eyeball this: we therefore proceed to quantify the impact of directional gain errors for different sources in the field. To do this, we begin by making the approximation that they are a function of distance from the calibrator source only, rather than specifically a function of direction and distance. This is simply because we don't have enough calibrators to do otherwise. Secondly, we treat the effect of directional gain errors as \textit{increasing the noise} around a source compared to what the thermal noise is expected to do: this thermal noise is expected to increase as a function of distance from the phase centre because we apply a beam-correction during imaging (i.e. because we do not image apparent flux but integrated flux). \cref{fig.noiserms.distEGS} shows this effect in practice; note that the noise estimation around LBCS6 (3C295) is biased because we are still quite close to this extremely bright source, and thus likely have a large contribution from artefacts. This explains its outlier behaviour.
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/RMSvsDistFromEGS.png}
\caption{Thermal noise near LBCS sources as a function of distance from the calibrator.}
\label{fig.noiserms.distEGS}
\end{figure}

\pg
To estimate the prevalence of directional gain errors, we then calculate the rms around each LBCS source. This is plotted in \cref{fig.dderms.distEGS}. The RMS around each source is normalised by the source's integrated flux as calculated on the restored image: this is because we expect brighter images to show stronger artefacts associated to directional gain errors. It should be noted that LBCS8 is not deconvolved at all: as such, its RMS is effectively calculated on a corrupted dirty map of the source, rather than a residual map. This helps explains its outlier behaviour. This outlier aside, the directional gain rms seems fairly flat as a function of distance from 3C295, however - a very good sign, as it indicates that the directional gain errors are comparable across the primary beam. However, we do not take into account the increase in thermal rms as a function of distance from phase centre in this plot. We therefore look at the normalised RMS values for each LBCS source, shown in \cref{fig.rmsratio.distEGS}. The relationship remains flat, though LBCS5 now stands as a $3\sigma$ outlier. The blue line shows the mean value, excluding LBCS8; it lies at $y=1.08063981$. In other words, the average RMS induced by directional gain errors, when normalised by both the flux of a source for which it is calculated and the thermal noise, is very close to unity, and does not increase very much as distance from 3C295 increases. While it apparently increases in the direction of LBCS5, the EGS is not in this direction from 3C295, and so this is not a source of concern for our project of imaging the EGS using international LOFAR.
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/ArtefactRMSvsDistFrom3c295.png}
\caption{DDE noise near LBCS sources as a function of distance from the phase centre.}% Because 3C295 is so much brighter than all other sources ($\sim100$Jy; second-brightest LBCS source is $\sim2$Jy), its DDE noise was normalised by its integrated model flux.}
\label{fig.dderms.distEGS}
\end{figure}
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/NormArtefactRMSvsDistFrom3c295.png}
\caption{Normalised RMS values around each of the LBCS sources in the field as a function of distance from phase centre. Here, `off-source' means that the rms is calculated near the source, not on the source (this is because some LBCS sources were too faint to be deconvolved: they were therefore still included in the residuals, and would bias the rms calculation).}
\label{fig.rmsratio.distEGS}
\end{figure}

\pg
The most encouraging result shown above, however, is for LBCS4: this source appear to be only very slightly affected by directional gain errors. Although LBCS3 seems affected by directional gain errors, it is not so affected that it is smeared out completely: we can thus expect there to still be signal at that distance from the calibrator and phase centre. Since the EGS lies between these two calibrator sources and 3C295, this is extremely positive. Similarly, the fact that LBCS4 resolves into two separate sources (as does LBCS1, for that matter) would indicate that even neighbouring bright sources do resolve properly - this is not a rigorous test, but it is a fortunate serendipitous observation to be able to make.

### 2018-11-12 source/acknowledgement-paragraph.tex
asdasda

\pg
I'll begin with thanks to Cyril Tasse, Philippe Zarka and Oleg Smirnov. Though each will claim to have only done a small part of the work in guiding and teaching me over the course of this PhD, they have all contributed monumentally - it was not always easy, but it was certainly worthwhile.

\pg
Similarly, I owe an infinite debt to the post-doctoral fellows and colleagues I had the pleasure to work with and befriend. I would particularly like to thank Marcel Atemkeng, Julien Girard, Roger Deane, Kshitij Thorat, Trienko Grobler, Arun Aniyan and Alan Loh - both for putting up my with ridiculous imprecations to my computer terminal, and for helping me refine my understanding of my own work. Of course, I am just as grateful to have known and worked with many other fine people: Corentin Louis, Lucas Grosset, and Lisa Bardou in France; Theophilius Ansah-Narh, Kelachukwu Iheanetu, Ulrich Mbou Sob, Kwazi Mthembu and all the RATT Masters students during my time in South Africa.

@@

\pg
I am also extremely grateful to those teachers who first helped me develop an interest in science and mathematics: without Nasreddine Nouar, Katrina Kemper, and Judith Byrd, I would not be what I am today. I hope to be as good a teacher to my own students as you were to me.

>>>>

\pg
I'll begin with thanks to Cyril Tasse, Philippe Zarka and Oleg Smirnov. Though each will claim to have only done a small part of the work in guiding and teaching me over the course of this PhD, they have all contributed monumentally - it was not always easy, but it was certainly worthwhile. I would also like to thank all the members of my thesis defense, many of whom crossed borders to attend. I would particularly like to thank my examiners - Neal Jackson, Patrick Charlot and Ciara Ferrari - whose thorough comments and invaluable insights greatly improved the quality of this work.

\pg
Similarly, I owe an infinite debt to the post-doctoral fellows and colleagues I had the pleasure to work with and befriend. I would particularly like to thank Marcel Atemkeng, Julien Girard, Roger Deane, Kshitij Thorat, Trienko Grobler, Arun Aniyan and Alan Loh - both for putting up with my ridiculous imprecations to my computer terminal, and for helping me refine my understanding of my own work. Of course, I am just as grateful to have known and worked with many other fine people: Corentin Louis, Lucas Grosset, and Lisa Bardou in France; Theophilius Ansah-Narh, Kelachukwu Iheanetu, Benjamin Hugo, Ulrich Mbou Sob, Kwazi Mthembu and all the RATT Masters students during my time in South Africa.

@@

\pg
I am also extremely grateful to those teachers who first helped me develop an interest in science and mathematics: without Nasreddine Nouar, Katrina Kemper, and Judith Byrd, I would not be what I am today. I hope to be as good a teacher to my own students as you were to me. Similarly, I would like to thank those who encouraged me to pursue this interest in higher education: Catherine Heymans, Andreas Zech and Jacques le Bourlot. Yours is an example I strive to emulate.
